---
title: "Coursera Regression Models Project"
author: "Samir N. Hag Ibrahim"
date: "7/10/2020"
output:
  html_document:
    df_print: paged
---
## Introduction
The magazine known as Motor Trend (MT) covers the automobile industry with an interest in seeing _the relationship of MPG (miles per gallan) and a number of variables_. The following questions need to be considered:

   1- “Is an automatic or manual transmission better for MPG ?”
   2- “Quantify the MPG difference between automatic and manual transmissions”

## 1- Data loading and Exploration
```{r}
library(tidyverse)
data(mtcars)
head(mtcars)
```

exploring the data structure.

```{r}
str(mtcars)
```

So, we have 11 variable with 32 observation. Now checking for any missing entries in the data set.

```{r}
sum(is.na(mtcars))
```


Checking variable _am_ .

```{r}
unique(mtcars$am)
```

now the data is tidy and redy to EA.


## 2- Exploratory Analysis
checking the data summary

```{r}
summary(mtcars)
```

Now lets visualize variables' relationship.

```{r}
library(Hmisc)
library(PerformanceAnalytics)
chart.Correlation(mtcars, histogram = FALSE, pch = 12)
cm <- rcorr(as.matrix(mtcars))

cm
```

now lets convert _am_ into two-level factor variable

```{r}
mtcars$am <- factor(mtcars$am)
levels(mtcars$am) <- c("Auto", "Manual")
levels(mtcars$am)
str(mtcars)
with(mtcars, table(am))
```
from the summary results we can see that `r summary(mtcars$am)[1]` of the cars are _Automatic_ and `r summary(mtcars$am)[2]` of the cars are _Mannual_. 


## 3- Exploring the relationship btw MPG Vs am ("Auto", "MAnual")

lets remind ourselves the 1st question needs to be answered:
1- “Is an automatic or manual transmission better for MPG ?”

lest's figure out what kind of relationship do we have here, if any!.

```{r}
ggplot(mtcars, aes(am, mpg))+
        geom_boxplot(aes(fill= am))+
        labs(x = "Transmission", y = "Miles per Hour - mpg")
```

```{r}
means <- aggregate(mpg~am, mtcars, mean)
means
```
Appearently, Manula cars seems to have better mpg compared with automatic cars according to there average mpgs. now let's chack whether this difference is sig. or not.

```{r}
auto_cars <- subset(mtcars, am == "Auto") ; nrow(auto_cars) 
manual_cars <- subset(mtcars, am =="Manual"); nrow(manual_cars)
# nrow() just to check the subset is correct
```
now we have divided our dataset into two sets based on type of transmission
now we carry out t-test to see whether there is a sig. difference between to two data sets or not.

```{r}
ttest <- t.test(auto_cars$mpg, manual_cars$mpg)
ttest
```

there is a sig. difference with fixing other variables. Hence, we can say that _Manual cars is better for the mpgs._

## 4- Quantifying the relationships (MPG Vs "Auto" $ "Manual")
Now the second question was:
“Quantify the MPG difference between automatic and manual transmissions”

in order to quantify the relationship we needs to perform linear regression analysis

```{r}
fit1 <- lm(mpg~am, mtcars)
summary(fit1)
```
from the SLR we can see that, the slope for the Automatic cars is `r round (summary(fit1)$coef[1,1], 3)`, where for manual cars is `r round(summary(fit1)$coef[2,1], 3)` which is `r round(summary(fit1)$coef[2,1] - summary(fit1)$coef[1,1], 3)`more. on the other hand, $R^2$ equals 0.338 i.e. only 34% of the variations in the data were explained by _am_ as a variable. to be more realistic, we need to include other varaibles in the equation in order to explain the real relationship.


## 5- Multiregression Analysis
now we will conduct mulit regression analysis in order to get the best fit for the model and the more realistic relationship between transmissions and mpg.

```{r}
fit2 <- step(lm(mpg~., mtcars),direction = "both" ,trace = 0)
summary(fit2)
```
from multiregression anaylsis, it is clear that the model is better fitting the data now with $R^2$ = 0.834 (i.e. explaining 84% of the variation in the data). Also, the only significant variables for mpg are weight "wt", acceleration "qsec" and finally Manual transmission "amManual". and finally, Manual transmission has `r summary(fit2)$Coef[4,1]` mpg higer than that for the Automatic transmission.

## comparing model diagnostics

### Diagnostics for SLR (fit1)
```{r}
par(mfrow = c(2,2))
plot(fit1)
```

### Diagnostics for Multiregression model (fit2)
```{r}
par(mfrow = c(2,2))
plot(fit2)
```


























